P4IRS: An Intermediate Representation and Compiler for Parallel and Performance-Portable Particle Simulations

### 2023-11-19 manuscript.tex
Add more text for P4IRS section
Signed-off-by: Rafael Ravedutti <rafael.r.ravedutti@fau.de>

\autoref{lst:pairslj} shows a simple Lennard-Jones kernel implemented in P4IRS.
Functions can be provided through a method, and the parameters in the method refer to different particles in the system.
In this way, many-body potentials can also be described with methods with more than two parameters.
Functions that only updates every particle separately can also be described with a single parameter.
\autoref{lst:euler} displays how to implement a simple Euler integrator for \ac{MD}.

@@

@@

@@

>>>>

\autoref{lst:pairslj} shows a simple Lennard-Jones kernel implemented in P4IRS.
Functions can be provided through a method (line 1), and the parameters in the method refer to different particles in the system.
In this way, many-body potentials can also be described using methods with more than two parameters.
Some keywords are also provided by P4IRS for writing the kernels, in line 2 the \emph{squared\_distance} method is used for obtaining the squared distance for the atoms \emph{i} and \emph{j}, and in line 4 the \emph{delta} method is used for obtaining the delta vector.
Also, in lines 3 and 4 it is possible to see accesses to feature properties called \emph{sigma6} and \emph{epsilon}, which are defined during the simulation setup as properties that depends on the atom types feature.
In line 4, the \emph{apply} method is used for applying the computed term into the \emph{force} property of the particles, the method allows for identifying particle properties reductions and is also used when generating the kernel variant using the half neighbor-lists.

@@

Functions that update every particle separately can also be described with a single parameter.
\autoref{lst:euler} displays how to implement a simple Euler integrator for \ac{MD} (see line 1).
In lines 2-3, there are also property accesses for the particles' linear velocities, forces, masses and positions.
Note that vector operations are intrinsically supported in P4IRS, which is also the case for matrices and quaternions data structures.
P4IRS methods also allow the specification of parameters such as the \emph{dt} specified in line 3, these must be given when setting up the simulation.

@@

\autoref{lst:pairs_md_setup} shows the simulation setup for the \ac{MD} case. In the first part, a \emph{simulation} instance is created, and its identifier, the list of particle shapes used, the number of time-steps and the floating-point precision are specified.
Further, the properties needed for each particle in the simulation are specified, together with their data types and (if needed) initial values.
Some properties are also defined as volatile (in this case the \emph{force}) property, which means they are reset after every time-step.
To add feature properties, it is first required to add the specified feature (in this example, each atom has a feature named \emph{type}), and then specifiy the feature name for these properties.
Next, the simulation domain is specified and the initial state of the system is defined by reading data from a file, a list of properties must be given in the same order as they appear in the file, and the particles shape must also be included.

@@

After the properties and simulation settings, the optimization settings must also be specified (whether neighbor-lists are used, the reneighbouring frequency and if these must be half-neighbor lists).
The proper kernels are specified using the \emph{compute} method. The cutoff radius must be specified for kernels with more than one parameter/particle and other parameters used in such kernels (such as the \emph{dt} in the Euler kernel) must also be provided as a Python dictionary.
